The polynomial representation of the type $A_{n - 1}$ rational Cherednik algebra in characteristic $p \mid n$

### 2015-05-19 formalwriteup.tex
Add commentary for CI

Let the vectors $f_i \in A$ generate an ideal $I \subset A$.  We will now show that $A/I$ is a complete intersection.  For this, we require some preparatory computations.

\subsection{Derivative computations}

@@

\todo[inline]{Split computations out into some lemmas on computations of $F'(z)$ and $V(z)$ which explicitly mention the hypotheses that $\sum_i m_i y_i = 0$.}

To show that $A/I$ is a complete intersection, we will show that if $x_0,\dots,x_{n-1} \in k$ satisfy $f_i(x_0,\dots,x_{n-1})=0$ for all $i$, then $x_0=\dots=x_{n-1}=0$. We will use a `residues by parts' argument; using the fact that the residue of an exact differential is zero, we will relate the coefficients of a power series to the coefficients of the derivative of that power series and derive a contradiction. To show this we must consider the derivatives of some power series in $z$ (including $F(z)$) with coefficients from $k$ rather than coefficients from $A$.

\begin{lemma}\label{lem:dFdz2}
For any $y_0,\dots,y_{r-1}$ distinct elements of $k$ with $m_i$ positive integers such that $\sum_i m_iy_i=0$, if we let $g(z)=\prod_{i=0}^{r-1} (1-y_iz)^{m_i}$ and $F(z)=\sum_{k=0}^{p-1} \binom{c}{k}(g-1)^k$, then
\[
F'(z)=V(z)+\left(\sum_j \frac{-cm_jy_j}{1-y_jz}F(z)\right),
\]
where $[z^l]V(z)=0$ for $0 \le l \le p-1$. 
\end{lemma}
\begin{proof}
The computation proceeds essentially as it did for Lemma~\ref{lem:dFdz}.

We see easily that 
\[
\frac{\partial g}{\partial z} = g(z) \sum_j \frac{-m_jy_j}{1-y_jz}.
\]
We now consider $\frac{\partial F}{\partial z}$. We compute
\begin{align*}
\frac{\partial F}{\partial z}%&=\frac{\partial}{\partial z}\left(\sum_{k=0}^{p-1} \binom{c}{k} (g-1)^k\right)\\
&=\left(\sum_{k=1}^{p-1}k\binom{c}{k}(g-1)^{k-1}\right)\frac{\partial g}{\partial z}\\
%&=\left(\sum_j \frac{-x_j}{1-x_jz}\right)\left(\sum_{k=1}^{p-1}k\binom{c}{k}(g-1)^{k-1}\right)g(z)\\
%&=\left(\sum_j \frac{-x_j}{1-x_jz}\right)\left(\sum_{k=0}^{p-2}(k+1)\binom{c}{k+1}(g-1)^{k}\right)g(z)\\
&=\left(\sum_j\frac{-m_jy_j}{1-y_jz}\right)\left(\sum_{k=0}^{p-2}c\binom{c-1}{k}(g-1)^{k}\right)(g-1+1)\\
%&=\left(\sum_j \frac{-x_j}{1-x_jz}\right)\left(\left(\sum_{k=0}^{p-2}c\binom{c-1}{k}(g-1)^{k}\right)+\left(\sum_{k=0}^{p-2}c\binom{c-1}{k}(g-1)^{k+1}\right)\right)\\
&=\left(\sum_j\frac{-m_jy_j}{1-y_jz}\right)\left(\left(\sum_{k=0}^{p-2}c\binom{c-1}{k}(g-1)^{k}\right)+\left(\sum_{k=1}^{p-1}c\binom{c-1}{k-1}(g-1)^{k}\right)\right)\\
&=\left(\sum_j \frac{-m_jy_j}{1-y_jz}\right)\left(-c\binom{c-1}{p-1}(g-1)^{p-1}+\sum_{k=0}^{p-1}c\binom{c}{k}(g-1)^{k}\right)\\
&=\left(\sum_j \frac{-cm_jy_j}{1-y_jz}\right)F(z)+\left(\sum_j \frac{-m_jy_j}{1-y_jz}\right)\left(-c\binom{c-1}{p-1}(g-1)^{p-1}\right).
\end{align*}
We define 
\[
V(z)=\left(\sum_j \frac{-m_jy_j}{1-y_jz}\right)\left(-c\binom{c-1}{p-1}(g-1)^{p-1}\right).
\]
It is then clear that 
\[
F'(z)=V(z)+\left(\sum_j \frac{-cm_jy_j}{1-y_jz}F(z)\right).
\]
Then it remains only to show that $[z^l]V(z)=0$ for $l=0,\dots,p-1$. To show this, we need only show that $V(z)=z^pv(z)$ for some power series $v(z)$.

We consider the expansion of $g(z)$; this is $g(z)=\prod_j (1-y_jz)^{m_j} = 1-z\left(\sum_i m_iy_i \right)+z^2e_2(y_i)+\dots+z^ne_n(y_i)$ where $e_2,\dots,e_n$ are polynomials in the $y_i$. In particular we note that $\sum_i m_iy_i=0$ by hypothesis, so $z^2 \mid g(z)-1$.

\begin{lemma} \label{lem:resbyparts}
Let $y_0,\dots,y_{r-1}$ be distinct non-zero elements of $k$ and $m_i$ positive integers so that that $\sum_i m_iy_i=0$.  For $F(z)$ as in Lemma \ref{lem:dFdz2}, let $a(z)$ be the formal power series
\[
a(z)=\frac{1}{\prod_{i=0}^{r-1}(1-y_iz)}F(z).
\]
For any integer $l \le p-r$, $[z^l]a(z)$ is a linear combination of $[z^{l+1}]a(z)$, $\ldots$, $[z^{l+r}]a(z)$.
\end{lemma}
\begin{proof} We begin by showing that 
\[
a'(z)=z^pb(z)-\sum_j \frac{(m_jc-1)y_j}{1-y_jz}a(z)
\]
for some formal power series $b(z)$. By Lemma~\ref{lem:dFdz2} we have
\[
F'(z)=V(z)+\left(\sum_j \frac{-cm_jy_j}{1-y_jz}F(z)\right),
\]
where $[z^l]V(z)=0$ for $0 \le l \le p-1$. 
We now compute
\begin{align*}
a'(z)&=\frac{d}{dz}\left(\frac{1}{\prod_{i=0}^{r-1}(1-y_iz)}F(z)\right)\\
&=\frac{1}{\prod_{i=0}^{r-1}(1-y_iz)}F'(z)+\sum_{j=0}^{r-1}\frac{y_j}{(1-y_jz)\prod_{i=0}^{r-1}(1-y_iz)}F(z)\\
&=\frac{1}{\prod_{i=0}^{r-1}(1-y_iz)}\left(V(z)+\sum_j \frac{-cm_jy_j}{1-y_jz}F(z)\right)+\sum_{j=0}^{r-1}\frac{y_j}{1-y_jz}a(z)\\
&=\frac{V(z)}{\prod_{i=0}^{r-1}(1-y_iz)}+\sum_j \frac{-cm_jy_j}{1-y_jz}a(z)+\sum_{j=0}^{r-1}\frac{y_j}{1-y_jz}a(z)\\
&=\frac{V(z)}{\prod_{i=0}^{r-1}(1-y_iz)}-\sum_j \frac{(m_jc-1)y_j}{1-y_jz}a(z).
\end{align*}
Since $[z^l]V(z)=0$ for $0 \le l \le p-1$ we see that 
\[
\frac{V(z)}{\prod_{i=0}^{r-1}(1-y_iz)}=z^pb(z)
\]
for some power series $b(z)$, giving the claim.

Now, for any integer $l$, define the Laurent polynomial 
\[
h_l(z)=z^{-l-r}\prod_i(1-y_iz).
\]
Notice that
\begin{align*}
\frac{d}{dz}( h_l(z) a(z)) &= -(l + r)z^{-1} h_l(z) a(z) - \sum_i \frac{y_i}{1 - y_iz} h_l(z) a(z) + h_l(z) a'(z) \\
&=  -(l + r)z^{-1} h_l(z) a(z) - \sum_i \frac{y_i}{1 - y_iz} h_l(z) a(z) - h_l(z) \sum_j \frac{(m_j c - 1)y_i}{1 - y_ijz} a(z)  + z^ph_l(z) b(z)\\
&= -(l + r)z^{-1} h_l(z) a(z) - \sum_i \frac{y_im_i c}{1 - y_iz} h_l(z) a(z) + z^ph_l(z) b(z)\\
&= -\left((l + r) z^{-1} + \sum_i \frac{y_im_i c}{1 - y_iz}\right) h_l(z) a(z) + z^ph_l(z) b(z).
\end{align*}
We note that 
\[
-\left((l + r) z^{-1} + \sum_i \frac{y_im_i c}{1 - y_iz}\right) h_l(z)
\]
is a Laurent polynomial with lowest degree term 
\[
- z^{-l-r-1} \left(l + r + \sum_i y_i m_i c\right) = - z^{-l-r-1}(l+ r)
\]
and highest degree term 
\[
- z^{-l - 1} (-1)^r  \prod_i y_i\left((l + r) - \sum_i m_i c\right).
\]
The lowest degree term of the formal power series $z^ph_l(z)b(z)$ has degree at least $p-l-r$. We note that we must have $[z^{-1}]\left(\frac{d}{dz}( h_l(z) a(z))\right)=0$, so for $p-l-r \ge 0$, we see that $[z^{-1}]\left(\frac{d}{dz}( h_l(z) a(z))\right)=0$ is a linear combination of $[z^l]a(z),[z^{l+1}]a(z),\dots,[z^{l+r}]a(z)$ in which $[z^l]a(z)$ has nonzero coefficient 
\[
(-1)^r  \prod_i y_i\left((l + r) - \sum_i m_i c\right),
\]
which is nonzero for $c$ generic and $\prod_i y_i \ne 0$. Therefore $[z^l]a(z)$ is a linear combination of $[z^{l+1}]a(z)$, $\dots$, $[z^{l+r}]a(z)$ when $l \le p-r$. 
\end{proof}

@@

We are now ready to show that $A/I$ is a complete intersection.

>>>>

Consider the ideal
\[
I = \langle f_0, \ldots, f_{n - 2} \rangle \subset A
\]
generated by the $f_i$.  In this section, we will show that $A/I$ is a complete intersection, for which we require some preparatory computations.

\subsection{Computations of derivatives}

Our goal in this subsection is to establish Lemma \ref{lem:resbyparts}, which is a formal power series analogue of \cite[Lemma 3.4]{CE}.  Our method will be an adaptation of the ``residues by parts'' argument which appears there.  First, we require some direct computations of derivatives.

@@

\begin{lemma}\label{lem:dFdz2}
Let $y_0,\dots,y_{r-1}$ be distinct elements of $k$ and $m_i$ positive integers such that $\sum_i m_iy_i=0$.  Define 
\[
g(z)=\prod_{i=0}^{r-1} (1-y_iz)^{m_i} \text{  and } F(z)=\sum_{k=0}^{p-1} \binom{c}{k}(g-1)^k.
\]
For some $V(z)$ satisfying $[z^l]V(z)=0$ for $0 \le l \le p-1$, we have
\[
F'(z)=V(z) - \sum_j \frac{cm_jy_j}{1-y_jz}F(z).
\]
\end{lemma}
\begin{proof}
We compute as in the proof of Lemma~\ref{lem:dFdz}. We see easily that 
\[
\frac{\partial g}{\partial z} = g(z) \sum_j \frac{-m_jy_j}{1-y_jz}.
\]
We now consider $\frac{\partial F}{\partial z}$. We compute
\begin{align*}
\frac{\partial F}{\partial z}%&=\frac{\partial}{\partial z}\left(\sum_{k=0}^{p-1} \binom{c}{k} (g-1)^k\right)\\
&=\sum_{k=1}^{p-1}k\binom{c}{k}(g-1)^{k-1}\frac{\partial g}{\partial z}\\
%&=\left(\sum_j \frac{-x_j}{1-x_jz}\right)\left(\sum_{k=1}^{p-1}k\binom{c}{k}(g-1)^{k-1}\right)g(z)\\
%&=\left(\sum_j \frac{-x_j}{1-x_jz}\right)\left(\sum_{k=0}^{p-2}(k+1)\binom{c}{k+1}(g-1)^{k}\right)g(z)\\
&=- \sum_j\frac{m_jy_j}{1-y_jz}\left(\sum_{k=0}^{p-2}c\binom{c-1}{k}(g-1)^{k}\right)(g-1+1)\\
%&=\left(\sum_j \frac{-x_j}{1-x_jz}\right)\left(\left(\sum_{k=0}^{p-2}c\binom{c-1}{k}(g-1)^{k}\right)+\left(\sum_{k=0}^{p-2}c\binom{c-1}{k}(g-1)^{k+1}\right)\right)\\
&=-\sum_j\frac{m_jy_j}{1-y_jz}\left(\sum_{k=0}^{p-2}c\binom{c-1}{k}(g-1)^{k}+\sum_{k=1}^{p-1}c\binom{c-1}{k-1}(g-1)^{k}\right)\\
&=-\sum_j \frac{m_jy_j}{1-y_jz}\left(-c\binom{c-1}{p-1}(g-1)^{p-1}+\sum_{k=0}^{p-1}c\binom{c}{k}(g-1)^{k}\right)\\
&=-\sum_j \frac{cm_jy_j}{1-y_jz}F(z) + \sum_j \frac{m_jy_j}{1-y_jz}c\binom{c-1}{p-1}(g-1)^{p-1}.
\end{align*}
Defining the formal power series
\[
V(z)=\sum_j \frac{m_jy_j}{1-y_jz} c\binom{c-1}{p-1}(g-1)^{p-1},
\]
we have shown that 
\[
F'(z)=V(z)- \sum_j \frac{cm_jy_j}{1-y_jz}F(z).
\]

It remains only to show that $[z^l]V(z)=0$ for $l=0,\dots,p-1$, for which it suffices to show that $z^p \mid V(z)$.  Consider the expansion
\[
g(z)=\prod_j (1-y_jz)^{m_j} = 1-z\left(\sum_i m_iy_i \right)+z^2e_2(y_0, \ldots, y_{r - 1})+\dots+z^ne_n(y_0, \ldots, y_{r - 1})
\]
for some polynomials $e_2,\dots,e_n$. Since $\sum_i m_iy_i=0$ by hypothesis, we obtain $z^2 \mid g(z)-1$.  By definition of $V(z)$, we see that $(g-1)^{p-1} \mid V(z)$, so $z^{2p-2} \mid V(z)$. Since $p \ge 2$, this implies $z^p \mid V(z)$, as needed.
\end{proof}

\begin{lemma} \label{lem:resbyparts}
Let $y_0,\dots,y_{r-1}$ be distinct non-zero elements of $k$ and $m_i$ positive integers so that that $\sum_i m_iy_i=0$.  For $F(z)$ as in Lemma \ref{lem:dFdz2}, let $a(z)$ be the formal power series
\[
a(z)=\frac{1}{\prod_{i=0}^{r-1}(1-y_iz)}F(z).
\]
For any integer $l \le p-r$, $[z^l]a(z)$ is a linear combination of $[z^{l+1}]a(z)$, $\ldots$, $[z^{l+r}]a(z)$.
\end{lemma}
\begin{proof} 
We will use an analogue of ``residues by parts'' in a formal power series setting.   We begin by showing that 
\[
a'(z)=z^pb(z)-\sum_j \frac{(m_jc-1)y_j}{1-y_jz}a(z)
\]
for some formal power series $b(z)$. By Lemma~\ref{lem:dFdz2} we have
\[
F'(z)=V(z)+\left(\sum_j \frac{-cm_jy_j}{1-y_jz}F(z)\right),
\]
where $[z^l]V(z)=0$ for $0 \le l \le p-1$. 
We now compute
\begin{align*}
a'(z)&=\frac{d}{dz}\left(\frac{1}{\prod_{i=0}^{r-1}(1-y_iz)}F(z)\right)\\
&=\frac{1}{\prod_{i=0}^{r-1}(1-y_iz)}F'(z)+\sum_{j=0}^{r-1}\frac{y_j}{(1-y_jz)\prod_{i=0}^{r-1}(1-y_iz)}F(z)\\
&=\frac{1}{\prod_{i=0}^{r-1}(1-y_iz)}\left(V(z)+\sum_j \frac{-cm_jy_j}{1-y_jz}F(z)\right)+\sum_{j=0}^{r-1}\frac{y_j}{1-y_jz}a(z)\\
&=\frac{V(z)}{\prod_{i=0}^{r-1}(1-y_iz)}+\sum_j \frac{-cm_jy_j}{1-y_jz}a(z)+\sum_{j=0}^{r-1}\frac{y_j}{1-y_jz}a(z)\\
&=\frac{V(z)}{\prod_{i=0}^{r-1}(1-y_iz)}-\sum_j \frac{(m_jc-1)y_j}{1-y_jz}a(z).
\end{align*}
Since $[z^l]V(z)=0$ for $0 \le l \le p-1$ we see that 
\[
\frac{V(z)}{\prod_{i=0}^{r-1}(1-y_iz)}=z^pb(z)
\]
for some power series $b(z)$, giving the claim.

Now, for any integer $l$, define the Laurent polynomial 
\[
h_l(z)=z^{-l-r}\prod_i(1-y_iz).
\]
Notice that
\begin{align*}
\frac{d}{dz}( h_l(z) a(z)) &= -(l + r)z^{-1} h_l(z) a(z) - \sum_i \frac{y_i}{1 - y_iz} h_l(z) a(z) + h_l(z) a'(z) \\
&=  -(l + r)z^{-1} h_l(z) a(z) - \sum_i \frac{y_i}{1 - y_iz} h_l(z) a(z) - h_l(z) \sum_j \frac{(m_j c - 1)y_i}{1 - y_ijz} a(z)  + z^ph_l(z) b(z)\\
&= -(l + r)z^{-1} h_l(z) a(z) - \sum_i \frac{y_im_i c}{1 - y_iz} h_l(z) a(z) + z^ph_l(z) b(z)\\
&= -\left((l + r) z^{-1} + \sum_i \frac{y_im_i c}{1 - y_iz}\right) h_l(z) a(z) + z^ph_l(z) b(z).
\end{align*}
We note that 
\[
-\left((l + r) z^{-1} + \sum_i \frac{y_im_i c}{1 - y_iz}\right) h_l(z)
\]
is a Laurent polynomial with lowest degree term 
\[
- z^{-l-r-1} \left(l + r + \sum_i y_i m_i c\right) = - z^{-l-r-1}(l+ r)
\]
and highest degree term 
\[
- z^{-l - 1} (-1)^r  \prod_i y_i\left((l + r) - \sum_i m_i c\right).
\]
The lowest degree term of the formal power series $z^ph_l(z)b(z)$ has degree at least $p-l-r$. We always have 
\[
[z^{-1}]\left(\frac{d}{dz}( h_l(z) a(z))\right)=0,
\]
so for $p-l-r \ge 0$, we see that $[z^{-1}]\left(\frac{d}{dz}( h_l(z) a(z))\right)=0$ is a linear combination of $[z^l]a(z),[z^{l+1}]a(z),\dots,[z^{l+r}]a(z)$ in which $[z^l]a(z)$ has nonzero coefficient 
\[
(-1)^r  \prod_i y_i\left((l + r) - \sum_i m_i c\right),
\]
which is nonzero for $c$ generic and $\prod_i y_i \ne 0$. Therefore $[z^l]a(z)$ is a linear combination of $[z^{l+1}]a(z)$, $\dots$, $[z^{l+r}]a(z)$ when $l \le p-r$. 
\end{proof}

@@

We now show that $A/I$ is a complete intersection.  Our technique is a translation of the idea of the proof of \cite[Theorem 3.2]{CE} to the formal power series context.
